Real-space computation of $E$/$B$-mode maps I: Formalism and Compact Kernels

### 2018-05-22 articles/formalism/sections/real_space_formalism.tex
Split kernel image and moved MG/MB to right after they are introduced.

 CMB polarization is measured in terms of Stokes parameters, time averages of the linear polarization of the electric field along cartesian axes perpendicular to the line of sight.\footnote{Throughout we use the conventions of HEALPix \cite{healpix_primer}, measuring the polarization angle East of South.} Thus Stokes $Q$ and $U$ depend on the choice of the local coordinate system, and a rotation by an angle $\psi$ around the line of sight transforms them as
%
\beq \label{eq:qu-rot}
\fqu' = \begin{bmatrix} \cos{2 \phi} &  \sin{2 \phi} \\ -\sin{2\phi} & \cos{2 \phi} \end{bmatrix} \fqu \,.
\eeq
Equivalently the object $_{\pm 2}\bar{X}(\hat{n}) = Q(\hat{n}) \pm i U (\hat{n})$ transforms as ${}_{\pm 2}f' = e^{\mp 2i\phi} {}_{s}f$ and hence forms a spin-2 field \cite{Zaldarriaga1997}.

@@

Rotating the cartesian coordinates in the tangent plane at location $\hat{n}_q$ by an angle $\phi$ about the local $\hat{z}_q$ axis, the Stokes parameters in the new coordinate system relate to those in the original coordinate system as:
$\mathcal{R}_{\hat{z}_q}(\phi)[{}_{+2}X(\hat{n}_q)] =  {}_{+2}X(\hat{n}_q) e^{-i2\phi} $.
This same rotation by $\phi$ alters the Euler angle $\alpha_{qe}$ (that aligns the $x$-axis at $\hat{n}_q$ along the geodesic to the location $\hat{n}_e$): $\mathcal{R}_{\hat{z}_q}(\phi)[\alpha_{qe}] = \alpha_{qe} - \phi$.  Therefore one can see that $\mathcal{R}_{\hat{z}_q}(\phi)[e^{-i2\alpha_{qe}}] =  e^{-i2\alpha_{qe}} e^{i2\phi}$.

Given these transformation properties, the combination ${}_{+2}X(\hat{n}_q)e^{-i2\alpha_{qe}}$ is invariant under rotations and must be spin-0 by definition:
\beq
\mathcal{R}_{\hat{z}_q}(\phi)[{}_{+2}X(\hat{n}_q)e^{-i2\alpha'_{qe}}] = {}_{+2}X(\hat{n}_q) e^{-i2\alpha_{qe}}
\eeq
\revisit{I'm having some trouble understanding the next 2 sentences.} The real part of the function is constructed by product of functions $(Q\cos{2\alpha}, U\sin{2 \alpha})$ having the same parity and hence the real part must have even parity.  At the same time, the imaginary part of the function is constructed by multiplying functions $(Q\sin{2 \alpha}, U\cos{2\alpha})$ of opposite parity and hence must have an odd parity. Therefore we can make the association that contributions to  $(E+iB)(\hat n_e)$ must be proportional to $ {}_{+2}X(\hat{n}_q) e^{-i2\alpha_{qe}}$.

The same rotation $\mathcal{R}_{\hat{z}_q}(\phi)$ leaves the Euler angle $|\beta_{qe}|$ unaltered (it measures the angular distance between the points).  Thus we can conclude that the contribution to $(E+iB)(\hat n_e)$ from the position $n_q$ must have the form 
\beq
{}_{+2}X(\hat{n}_q) f(\beta_{qe})  e^{-i2\alpha_{qe}}
\eeq
for some real function $f$.  Note that when the two locations coincide ($\beta_{qe}=0$) then  $\alpha_{qe}=0,2\pi,4\pi,\dots$, implying $E + iB \propto Q+iU$.  This is a contradiction because $Q+iU$ does not transform as a spin-0 field under local rotations, and so we must have $f(\beta_{qe} = 0 ) = 0$.   Hence the $E/B$ fields are necessarily non-local.  A similar contradiction arises when the two locations are diametrically opposite, $\beta_{qe} = \pi$, and so $f(\beta_{qe} = \pi ) = 0$.  Any such function $f$ will let us construct $E/B$-like scalar fields.  Below we derive the particular one that gives rise to our familiar $E/B$ modes.

@@

The standard construction of $E/B$ fields depend on the spin-raising and -lowering operators, and is usually carried out in harmonic space.  The spin-raising operator ($\eth$), applied to a field of spin-s $_{s}g$, results in a fields with spin-$(s+1)$: $(\eth _{s}g)' = e^{-i(s+1)\phi}(\eth _{s}g)$  \cite{goldberg67}.  The complementary spin-lowering operator $(\bar{\eth})$  similarly results with $(\bar{\eth} _{s}g)' = e^{-i(s-1)\phi}(\bar{\eth} _{s}g)$.  The complex spin-0 scalar now arise from  the spin-2 fields ${_{\pm 2}X}$ as follows,
%
\begin{subequations}\label{eq:ebdef}
\beqry
\mathcal{E}(\hat{n}) + i \mathcal{B}(\hat{n}) &=& -\bar{\eth}^2 _{+ 2}\bar{X}(\hat{n}) \,,\label{eq:ebdef_lower}\\
\mathcal{E}(\hat{n}) - i \mathcal{B}(\hat{n}) &=& -{\eth}^2 _{-2}\bar{X}(\hat{n}) \,.
\eeqry
\end{subequations}
%
%which by the virtue of being scalars are independent of coordinate definitions. %The real part of this complex spin-0 scalar field corresponds to the E-mode while the imaginary part to the B-modes \cite{Kamionkowski1997}. 
The $\cal E/B$ fields are defined locally at point $\hat n$ in terms of the derivative operators $\eth$ and $\bar \eth$.

@@

We now define the different polarization data vectors and their representation in real  space as and harmonic as follows,\footnote{We adopt a convention in which real space quantities are denoted by bar-ed variable while those in harmonic space are denoted by tilde-ed variables.}
%
\beqrys
\bar{S} &=& \bmat E \\ B  \emat_{2 \rm N_{\rm pix} \times 1} ~~~~;~~ \bar{X} = \bmat _{+2}X \\ _{-2}X \emat_{2 \rm N_{\rm pix} \times 1} ~~;~~\bar{P} =\fqu_{\tiny {2 \rm N_{\rm pix} \times 1}} \,, \\
\tilde{S} &=& \bmat a^{E} \\ a^{B} \emat _{2 \rm N_{\rm alms} \times 1}  ~~; ~~ \tilde{X} = \bmat _{+2} \tilde{X} \\ _{-2} \tilde{X} \emat_{2 \rm N_{\rm alms} \times 1} \,.
\eeqrys
%
The different symbols have the same meaning as that discussed in \sec{sec:pol-primer}, except that the subscript ${\ell m}$ for the spherical harmonic coefficients is suppressed for cleaner notation.

We define transformations between different representations of the polarization field (i.e. from $Q,U$ to $_{\pm2} X$):
%
\beqrys
\bar T &=& \qutox_{2 \rm N_{\rm pix} \times 2 \rm N_{\rm pix}} ~~;~~ \bar T^{-1} = \frac{1}{2} \bar T^{\dagger} \,, \\
\tilde T &=& -\qutox_{2 \rm N_{\rm alms} \times 2 \rm N_{\rm alms}} ~~;~~ \tilde T^{-1} = \frac{1}{2} \tilde T^{\dagger} \,,
\eeqrys
%
The sign conventions we have chosen match HEALPix.
Using the data vectors and the matrix operators defined above we can now express, in compact notation, the forward and inverse relations between different representations of the polarization data vectors as follows,
%
\begin{subequations} \label{eq:pol_data_relns}
  %Kevin's version
  \beqry
  \bar{X} &= \bar T  \bar{P}; &\qquad \bar{P} = \frac{1}{2} \bar T^{\dagger}  \bar{X} ; \\
  \tilde{X} &= \tilde T \tilde{S}; &\qquad \tilde{S} = \frac{1}{2}\tilde T^{\dagger} \tilde{X} \,.
  \eeqry
  Meanwhile the spherical harmonic transforms are written as:
  \beqry
  \bar X &=  {{}_2\mathcal{Y}}  \tilde X; &\qquad \tilde X ={{}_2\mathcal{Y}}^{\dagger}  \bar X  ; \\
  \bar S &=  {{}_0\mathcal{Y}} \tilde S; &\qquad  \tilde S =  {{}_0\mathcal{Y}}^{\dagger} \bar S \,.
  \eeqry
% Aditya's version
%\beqry
%\bar{X} &=& \bar T * \bar{P} ~~;~~\bar{P} = \frac{1}{2} \bar T^{\dagger} * \bar{X} \,, \\
%\bar X &=&  {{}_2\mathcal{Y}} * \tilde X  ~~;~~ \tilde X ={{}_2\mathcal{Y}}^{\dagger} * \bar X  \,, \\
%\tilde{X} &=& \tilde T * \tilde{S} ~~;~~ \tilde{S} = \frac{1}{2}\tilde T^{\dagger} * \tilde{X} \,.\\ 
%\bar S &=&  {{}_0\mathcal{Y}} * \tilde S ~~;~~  \tilde S =  {{}_0\mathcal{Y}}^{\dagger} * \bar S \,.
%\tilde X &=&  {{}_2\mathcal{Y}}^{\dagger} * \bar X ~~;~~ \tilde{X} = \tilde T * \tilde{S} \,, \\
%\bar{S} &=& {{}_0\mathcal{Y}}*\tilde S ~~;~~ \tilde{S} = \frac{1}{2}\tilde T^{\dagger} * \tilde{X} \,.\\
%\bar{X} &=& \bar T * \bar{P} ~~;~~ \tilde{X} = \tilde T * \tilde{S} \,, \\
%\bar{P} &=& \frac{1}{2} \bar T^{\dagger} * \bar{X} ~~;~~ \tilde{S} = \frac{1}{2}\tilde T^{\dagger} * \tilde{X} \,. \\
%\eeqry
\end{subequations}
%
Finally we introduce the operators that project harmonic space data vector to the $E$ or $B$ subspace,
%
\begin{subequations} \label{eq:har_eb_op}
\beqry
\tilde O_E &=& \bmat \mathbb{1} & \mathbb{0} \\ \mathbb{0} & \mathbb{0} \emat _{2 \rm N_{\rm alms} \times 2 \rm N_{\rm alms} }   ~~;~~ \tilde S_E = \tilde O_E  \tilde S ,\\
\tilde O_B &=& \bmat \mathbb{0} & \mathbb{0} \\ \mathbb{0} & \mathbb{1} \emat _{2 \rm N_{\rm alms} \times 2 \rm N_{\rm alms} } ~~; ~~ \tilde S_B = \tilde O_B  \tilde S .
\eeqry
\end{subequations}
%
Note that these harmonic space matrices are idempotent ($\tilde O_E  \tilde O_E = \tilde O_E;  \tilde O_B  \tilde O_B= \tilde O_B$), orthogonal ($\tilde O_E  \tilde O_B = \mathbb{0}$), and sum to the identity matrix ($\tilde O_E + \tilde O_B = \mathbb{1}$).
%
%\begin{subequations} \label{eq:har_op_prop}
%\beqry
%\tilde O_E  \tilde O_E&=& \tilde O_E ~~;~~  \tilde O_B  \tilde O_B= \tilde O_B \,,\\
% \tilde O_E  \tilde O_B&=& \mathbb{0} \,, \label{eq:op_eb_ortho}\\ 
% \tilde O_E + \tilde O_B&=& \mathbb{1} \,.
%\eeqry
%\end{subequations}
%
The above relations for these harmonic space operators are exactly valid.  In the following sections we aim to derive the real space analogues ($O_E,O_B$) of these harmonic space operators.

@@

%
\begin{figure}[!t]
\centering
\includegraphics[width=0.5\columnwidth]{radiation_convolution.pdf}
\caption{The local cartesian coordinate are only drawn on the red circle, representative of the coordinate dependence of the Stokes parameters. The value of the scalar field at location `e' can be evaluated by summing over the contribution from all the Stokes parameters on the red circle (sphere). The convolutions are performed with kernels which are defined in term of the Euler angle $\gamma_{eq}$. Alternatively, one can compute the contribution from the Stokes parameter at location `q' to all the point on the blue circle(sphere) and this is a function of the Euler angle $\alpha_{eq}$. In the flat sky limit, since $\gamma=-\alpha$ there is no difference between the radiating and convolving kernels.}
\label{fig:planar_euler_angles}
\end{figure}
%

@@

%--------------------------------------------------------
\subsection{Evaluating Stokes $Q$/$U$ from scalar $E$/$B$}\label{sec:eb2qu}
The real space operator which translates $E$ \& $B$ fields to Stokes parameters $Q$ \& $U$ can be derived using a similar procedure. Expressed in the matrix-vector notation, inverse operator is given by,
%
\begin{subequations}
\beqry
\bar{P} &=& \bar{T}^{-1} {{}_2\mathcal{Y}}\, \tilde T {{}_0\mathcal{Y}^{\dagger}}\bar{S} = \frac{1}{2} \bar{T}^{\dagger} {{}_2\mathcal{Y}} \,\tilde T {{}_0\mathcal{Y}^{\dagger}}\bar{S}\,,  \\
&=&  \bar O^{-1} \bar{S}\,.
\eeqry
\end{subequations}
%
%The inverse operator $\bar{O}^{-1}$ can be derived by realizing that $\mm^{-1}$ is given by the following expression,
%%
%\beq
%\mm^{-1} = \sum_{\ell m}  {{}_{2}}Y_{\ell m}(\hat n_q) {}_{0}Y^*_{\ell m}(\hat n_e) \, 
%\eeq
%%
The inverse operator expressed in terms of the function $\mm$ given in \eq{eq:qu2eb_gen_kernel} is given by the following equation,
%
\beq
{\bar O}^{-1}=-\bmat \mathcal{M}_{r} & -\mathcal{M}_{i} \\  \mathcal{M}_{i}  & \mathcal{M}_{r} \emat_{2 N_{\rm pix} \times 2 N_{pix}} =-{{}_{\mm}f}(\beta_{eq},\ell_{\rm min},\ell_{\rm max})\bmat \cos(2 \alpha_{qe}) & -\sin(2\alpha_{qe})\\  \sin(2 \alpha_{qe})  & \cos(2 \alpha_{qe}) \emat \,,
\eeq
%
where all the symbols have the same meaning as discussed in \sec{sec:qu2eb}. Note that the kernel in the above equation differs from the one in \eq{eq:op_qu2eb_rad} by a change in sign on the off-diagonals of the block matrix. When expressed in terms of the same set of Euler angles used to define the operator $\bar{O}$, it can be shown that the different forms of the real space operator are given by,
%
\beqry
    {}_{+2}X(\hat{n}_q) &=  \sum_{e=1}^{N_{\rm pix}} [E+iB](\hat{n}_{e})\   \mathcal{M}^*_{B}[\hat{n}_e] \hspace{0.8cm} &\textrm{\emph{Radiation kernel}}\,\\
    {}_{+2}X(\hat{n}_q) &= \Bigg\lbrace \mathcal{M}^*_{G} \star [E+iB] \Bigg\rbrace(\hat{n}_q) \hspace{0.8cm} &\textrm{\emph {Convolution kernel}}\,, \label{eq:eb2qu_convolution}
\eeqry
%
where all the symbols have the same meaning as defined in \sec{sec:qu2eb}. Note that the conjugated forms of the Green's function and the effective beam for the operator $\bar{O}$ have their roles reversed for the inverse operator $\bar{O}^{-1}$.
%--------------------------------------------------------

>>>>

 CMB polarization is measured in terms of Stokes parameters, time averages of the linear polarization of the electric field along cartesian axes perpendicular to the line of sight.\footnote{Throughout we use the conventions of HEALPix \cite{healpix_primer}, measuring the polarization angle East of South.} Thus Stokes $Q$ and $U$ depend on the choice of the local coordinate system, and a rotation by an angle $\psi$ around the line of sight transforms them as
%
\beq \label{eq:qu-rot}
\fqu' = \begin{bmatrix} \cos{2 \psi} &  \sin{2 \psi} \\ -\sin{2\psi} & \cos{2 \psi} \end{bmatrix} \fqu \,.
\eeq
Equivalently the object $_{\pm 2}\bar{X}(\hat{n}) = Q(\hat{n}) \pm i U (\hat{n})$ transforms as ${}_{\pm 2}f' = e^{\mp 2i\psi} {}_{s}f$ and hence forms a spin-2 field \cite{Zaldarriaga1997}.

@@

Rotating the cartesian coordinates in the tangent plane at location $\hat{n}_q$ by an angle $\psi$ about the local $\hat{z}_q$ axis, the Stokes parameters in the new coordinate system relate to those in the original coordinate system as:
$\mathcal{R}_{\hat{z}_q}(\psi)[{}_{+2}X(\hat{n}_q)] =  {}_{+2}X(\hat{n}_q) e^{-i2\psi} $.
This same rotation by $\psi$ alters the Euler angle $\alpha_{qe}$ (that aligns the $x$-axis at $\hat{n}_q$ along the geodesic to the location $\hat{n}_e$): $\mathcal{R}_{\hat{z}_q}(\psi)[\alpha_{qe}] = \alpha_{qe} - \psi$.  Therefore one can see that $\mathcal{R}_{\hat{z}_q}(\psi)[e^{-i2\alpha_{qe}}] =  e^{-i2\alpha_{qe}} e^{i2\psi}$.

Given these transformation properties, the combination ${}_{+2}X(\hat{n}_q)e^{-i2\alpha_{qe}}$ is invariant under rotations and must be spin-0 by definition:
\beq
\mathcal{R}_{\hat{z}_q}(\psi)[{}_{+2}X(\hat{n}_q)e^{-i2\alpha'_{qe}}] = {}_{+2}X(\hat{n}_q) e^{-i2\alpha_{qe}}
\eeq
\revisit{I'm having some trouble understanding the next 2 sentences.} The real part of the function is constructed by product of functions $(Q\cos{2\alpha}, U\sin{2 \alpha})$ having the same parity and hence the real part must have even parity.  At the same time, the imaginary part of the function is constructed by multiplying functions $(Q\sin{2 \alpha}, U\cos{2\alpha})$ of opposite parity and hence must have an odd parity. Therefore we can make the association that contributions to  $(E+iB)(\hat n_e)$ must be proportional to $ {}_{+2}X(\hat{n}_q) e^{-i2\alpha_{qe}}$.

The same rotation $\mathcal{R}_{\hat{z}_q}(\psi)$ leaves the Euler angle $|\beta_{qe}|$ unaltered (it measures the angular distance between the points).  Thus we can conclude that the contribution to $(E+iB)(\hat n_e)$ from the position $n_q$ must have the form 
\beq
{}_{+2}X(\hat{n}_q) f(\beta_{qe})  e^{-i2\alpha_{qe}}
\eeq
for some real function $f$.  Note that when the two locations coincide ($\beta_{qe}=0$) then  $\alpha_{qe}=0,2\pi,4\pi,\dots$, implying $E + iB \propto Q+iU$.  This is a contradiction because $Q+iU$ does not transform as a spin-0 field under local rotations, and so we must have $f(\beta_{qe} = 0 ) = 0$.   Hence the $E/B$ fields are necessarily non-local.  A similar contradiction arises when the two locations are diametrically opposite, $\beta_{qe} = \pi$, and so $f(\beta_{qe} = \pi ) = 0$.  Any such function $f$ will let us construct $E/B$-like scalar fields.  Below we derive the particular one that gives rise to our familiar $E/B$ modes.

@@

The standard construction of $E/B$ fields depend on the spin-raising and -lowering operators, and is usually carried out in harmonic space.  The spin-raising operator ($\eth$), applied to a field of spin-s $_{s}g$, results in a fields with spin-$(s+1)$: $(\eth _{s}g)' = e^{-i(s+1)\psi}(\eth _{s}g)$  \cite{goldberg67}.  The complementary spin-lowering operator $(\bar{\eth})$  similarly results with $(\bar{\eth} _{s}g)' = e^{-i(s-1)\psi}(\bar{\eth} _{s}g)$.  The complex spin-0 scalar now arise from  the spin-2 fields ${_{\pm 2}X}$ as follows,
%
\begin{subequations}\label{eq:ebdef}
\beqry
\mathcal{E}(\hat{n}) + i \mathcal{B}(\hat{n}) &=& -\bar{\eth}^2 _{+ 2}\bar{X}(\hat{n}) \,,\label{eq:ebdef_lower}\\
\mathcal{E}(\hat{n}) - i \mathcal{B}(\hat{n}) &=& -{\eth}^2 _{-2}\bar{X}(\hat{n}) \,.
\eeqry
\end{subequations}
%
%which by the virtue of being scalars are independent of coordinate definitions. %The real part of this complex spin-0 scalar field corresponds to the E-mode while the imaginary part to the B-modes \cite{Kamionkowski1997}. 
The $\cal E/B$ fields are defined locally at point $\hat n$ in terms of the derivative operators $\eth$ and $\bar \eth$.

@@

We now define the different polarization data vectors and their representation in real  space as and harmonic as follows,\footnote{We adopt a convention in which real space quantities are denoted by bar-ed variable while those in harmonic space are denoted by tilde-ed variables.}
%
\beqrys
\bar{S} &=& \bmat E \\ B  \emat_{2 \rm N_{\rm pix} \times 1};\qquad \bar{X} = \bmat _{+2}X \\ _{-2}X \emat_{2 \rm N_{\rm pix} \times 1};\qquad \bar{P} =\fqu_{\tiny {2 \rm N_{\rm pix} \times 1}} \,, \\
\tilde{S} &=& \bmat a^{E} \\ a^{B} \emat _{2 \rm N_{\rm alms} \times 1};\qquad \tilde{X} = \bmat _{+2} \tilde{X} \\ _{-2} \tilde{X} \emat_{2 \rm N_{\rm alms} \times 1} \,.
\eeqrys
%
The different symbols have the same meaning as that discussed in \sec{sec:pol-primer}, except that the subscript ${\ell m}$ for the spherical harmonic coefficients is suppressed for cleaner notation.

We define transformations between different representations of the polarization field (i.e. from $Q,U$ to $_{\pm2} X$):
%
\beqrys
\bar T &=& \qutox_{2 \rm N_{\rm pix} \times 2 \rm N_{\rm pix}} ;\qquad \bar T^{-1} = \frac{1}{2} \bar T^{\dagger} \,, \\
\tilde T &=& -\qutox_{2 \rm N_{\rm alms} \times 2 \rm N_{\rm alms}};\qquad \tilde T^{-1} = \frac{1}{2} \tilde T^{\dagger} \,,
\eeqrys
%
The sign conventions we have chosen match HEALPix.
Using the data vectors and the matrix operators defined above we can now express, in compact notation, the forward and inverse relations between different representations of the polarization data vectors as follows,
%
\begin{subequations} \label{eq:pol_data_relns}
  %Kevin's version
  \beqry
  \bar{X} &= \bar T  \bar{P}; &\qquad \bar{P} = \frac{1}{2} \bar T^{\dagger}  \bar{X} ; \\
  \tilde{X} &= \tilde T \tilde{S}; &\qquad \tilde{S} = \frac{1}{2}\tilde T^{\dagger} \tilde{X} \,.
  \eeqry
  Meanwhile the spherical harmonic transforms are written as:
  \beqry
  \bar X &=  {{}_2\mathcal{Y}}  \tilde X; &\qquad \tilde X ={{}_2\mathcal{Y}}^{\dagger}  \bar X  ; \\
  \bar S &=  {{}_0\mathcal{Y}} \tilde S; &\qquad  \tilde S =  {{}_0\mathcal{Y}}^{\dagger} \bar S \,.
  \eeqry
% Aditya's version
%\beqry
%\bar{X} &=& \bar T * \bar{P} ~~;~~\bar{P} = \frac{1}{2} \bar T^{\dagger} * \bar{X} \,, \\
%\bar X &=&  {{}_2\mathcal{Y}} * \tilde X  ~~;~~ \tilde X ={{}_2\mathcal{Y}}^{\dagger} * \bar X  \,, \\
%\tilde{X} &=& \tilde T * \tilde{S} ~~;~~ \tilde{S} = \frac{1}{2}\tilde T^{\dagger} * \tilde{X} \,.\\ 
%\bar S &=&  {{}_0\mathcal{Y}} * \tilde S ~~;~~  \tilde S =  {{}_0\mathcal{Y}}^{\dagger} * \bar S \,.
%\tilde X &=&  {{}_2\mathcal{Y}}^{\dagger} * \bar X ~~;~~ \tilde{X} = \tilde T * \tilde{S} \,, \\
%\bar{S} &=& {{}_0\mathcal{Y}}*\tilde S ~~;~~ \tilde{S} = \frac{1}{2}\tilde T^{\dagger} * \tilde{X} \,.\\
%\bar{X} &=& \bar T * \bar{P} ~~;~~ \tilde{X} = \tilde T * \tilde{S} \,, \\
%\bar{P} &=& \frac{1}{2} \bar T^{\dagger} * \bar{X} ~~;~~ \tilde{S} = \frac{1}{2}\tilde T^{\dagger} * \tilde{X} \,. \\
%\eeqry
\end{subequations}
%
Finally we introduce the operators that project harmonic space data vector to the $E$ or $B$ subspace,
%
\begin{subequations} \label{eq:har_eb_op}
\beqry
\tilde O_E &=& \bmat \mathbb{1} & \mathbb{0} \\ \mathbb{0} & \mathbb{0} \emat _{2 \rm N_{\rm alms} \times 2 \rm N_{\rm alms} }   ~~;~~ \tilde S_E = \tilde O_E  \tilde S ,\\
\tilde O_B &=& \bmat \mathbb{0} & \mathbb{0} \\ \mathbb{0} & \mathbb{1} \emat _{2 \rm N_{\rm alms} \times 2 \rm N_{\rm alms} } ~~; ~~ \tilde S_B = \tilde O_B  \tilde S .
\eeqry
\end{subequations}
%
Note that these harmonic space matrices are idempotent ($\tilde O_E  \tilde O_E = \tilde O_E;  \tilde O_B  \tilde O_B= \tilde O_B$), orthogonal ($\tilde O_E  \tilde O_B = \mathbb{0}$), and sum to the identity matrix ($\tilde O_E + \tilde O_B = \mathbb{1}$).
%
%\begin{subequations} \label{eq:har_op_prop}
%\beqry
%\tilde O_E  \tilde O_E&=& \tilde O_E ~~;~~  \tilde O_B  \tilde O_B= \tilde O_B \,,\\
% \tilde O_E  \tilde O_B&=& \mathbb{0} \,, \label{eq:op_eb_ortho}\\ 
% \tilde O_E + \tilde O_B&=& \mathbb{1} \,.
%\eeqry
%\end{subequations}
%
The above relations for these harmonic space operators are exactly valid.  In the following sections we aim to derive the real space analogues ($O_E,O_B$) of these harmonic space operators.

@@

%
\begin{figure}[!t]
\centering
\includegraphics[width=0.5\columnwidth]{radiation_convolution.pdf}
\caption{The local cartesian coordinate are only drawn on the red circle, representative of the coordinate dependence of the Stokes parameters. The value of the scalar field at location `e' can be evaluated by summing over the contribution from all the Stokes parameters on the red circle (sphere). The convolutions are performed with kernels which are defined in term of the Euler angle $\gamma_{eq}$. Alternatively, one can compute the contribution from the Stokes parameter at location `q' to all the point on the blue circle(sphere) and this is a function of the Euler angle $\alpha_{eq}$. In the flat sky limit, since $\gamma=-\alpha$ there is no difference between the radiation and convolution kernels.}
\label{fig:planar_euler_angles}
\end{figure}
%

@@

%--------------------------------------------------------
\subsection{Evaluating Stokes $Q$/$U$ from scalar $E$/$B$}\label{sec:eb2qu}
The real space operator which translates $E$ \& $B$ fields to Stokes parameters $Q$ \& $U$ can be derived using a similar procedure. Expressed in the matrix-vector notation, inverse operator is given by,
%
\begin{subequations}
\beqry
\bar{P} &=& \bar{T}^{-1} {{}_2\mathcal{Y}}\, \tilde T {{}_0\mathcal{Y}^{\dagger}}\bar{S} = \frac{1}{2} \bar{T}^{\dagger} {{}_2\mathcal{Y}} \,\tilde T {{}_0\mathcal{Y}^{\dagger}}\bar{S}\,,  \\
&=&  \bar O^{-1} \bar{S}\,.
\eeqry
\end{subequations}
%
%The inverse operator $\bar{O}^{-1}$ can be derived by realizing that $\mm^{-1}$ is given by the following expression,
%%
%\beq
%\mm^{-1} = \sum_{\ell m}  {{}_{2}}Y_{\ell m}(\hat n_q) {}_{0}Y^*_{\ell m}(\hat n_e) \, 
%\eeq
%%
The inverse operator expressed in terms of the function $\mm$ given in \eq{eq:qu2eb_gen_kernel} is given by the following equation,
%
\beq
{\bar O}^{-1}=-\bmat \mathcal{M}_{r} & -\mathcal{M}_{i} \\  \mathcal{M}_{i}  & \mathcal{M}_{r} \emat_{2 N_{\rm pix} \times 2 N_{pix}} =-{{}_{\mm}f}(\beta_{eq},\ell_{\rm min},\ell_{\rm max})\bmat \cos(2 \alpha_{qe}) & -\sin(2\alpha_{qe})\\  \sin(2 \alpha_{qe})  & \cos(2 \alpha_{qe}) \emat \,,
\eeq
%
where all the symbols have the same meaning as discussed in \sec{sec:qu2eb}. Note that the kernel in the above equation differs from the one in \eq{eq:op_qu2eb_rad} by a change in sign on the off-diagonals of the block matrix. When expressed in terms of the same set of Euler angles used to define the operator $\bar{O}$, it can be shown that the different forms of the real space operator are given by,
%
\beqry
    {}_{+2}X(\hat{n}_q) &=  \sum_{e=1}^{N_{\rm pix}} [E+iB](\hat{n}_{e})\   \mathcal{M}^*_{B}[\hat{n}_e] \hspace{0.8cm} &\textrm{\emph{Radiation kernel}}\,\\
    {}_{+2}X(\hat{n}_q) &= \Bigg\lbrace \mathcal{M}^*_{G} \star [E+iB] \Bigg\rbrace(\hat{n}_q) \hspace{0.8cm} &\textrm{\emph {Convolution kernel}}\,, \label{eq:eb2qu_convolution_explicit}
\eeqry
%
where all the symbols have the same meaning as defined in \sec{sec:qu2eb}. Note that the conjugated forms of the Green's function and the effective beam for the operator $\bar{O}$ have their roles reversed for the inverse operator $\bar{O}^{-1}$.
%--------------------------------------------------------

\subsection{Visualizing the real space kernels} \label{sec:visualize_operator}
%
\newlength{\kernelfigwidth}
\setlength{\kernelfigwidth}{0.2\columnwidth}
\newlength{\kernelfigspace}
\setlength{\kernelfigspace}{-1.8mm}

\begin{figure}[t]

  %%%% NEW MG/MB
  \begin{center}
    \begin{tabular}{m{8ex}m{\kernelfigwidth}m{\kernelfigwidth}|m{\kernelfigwidth}m{\kernelfigwidth}}
%      & \multicolumn{2}{c}{$E+iB$} & \multicolumn{2}{c}{$Q+iU$} \\
    $b=90^\circ$ &
\hspace{\kernelfigspace}\includegraphics[width=\kernelfigwidth]{new_kernel/qu2eb_rker_rad_lat90_lon45.pdf} &
\hspace{\kernelfigspace}\includegraphics[width=\kernelfigwidth]{new_kernel/qu2eb_iker_rad_lat90_lon45.pdf} &
\hspace{\kernelfigspace}\includegraphics[width=\kernelfigwidth]{new_kernel/qu2eb_rker_con_lat90_lon45.pdf} &
\hspace{\kernelfigspace}\includegraphics[width=\kernelfigwidth]{new_kernel/qu2eb_iker_con_lat90_lon45.pdf} \\
$b=87^\circ$&
\hspace{\kernelfigspace}\includegraphics[width=\kernelfigwidth]{new_kernel/qu2eb_rker_rad_lat87_lon45.pdf} &
\hspace{\kernelfigspace}\includegraphics[width=\kernelfigwidth]{new_kernel/qu2eb_iker_rad_lat87_lon45.pdf} &
\hspace{\kernelfigspace}\includegraphics[width=\kernelfigwidth]{new_kernel/qu2eb_rker_con_lat87_lon45.pdf} &
\hspace{\kernelfigspace}\includegraphics[width=\kernelfigwidth]{new_kernel/qu2eb_iker_con_lat87_lon45.pdf} \\
$b=80^\circ$&
\hspace{\kernelfigspace}\includegraphics[width=\kernelfigwidth]{new_kernel/qu2eb_rker_rad_lat80_lon30.pdf} &
\hspace{\kernelfigspace}\includegraphics[width=\kernelfigwidth]{new_kernel/qu2eb_iker_rad_lat80_lon30.pdf} &
\hspace{\kernelfigspace}\includegraphics[width=\kernelfigwidth]{new_kernel/qu2eb_rker_con_lat80_lon30.pdf} &
\hspace{\kernelfigspace}\includegraphics[width=\kernelfigwidth]{new_kernel/qu2eb_iker_con_lat80_lon30.pdf} \\
$b=0^\circ$&
\hspace{\kernelfigspace}\includegraphics[width=\kernelfigwidth]{new_kernel/qu2eb_rker_rad_lat0_lon90.pdf} &
\hspace{\kernelfigspace}\includegraphics[width=\kernelfigwidth]{new_kernel/qu2eb_iker_rad_lat0_lon90.pdf} &
\hspace{\kernelfigspace}\includegraphics[width=\kernelfigwidth]{new_kernel/qu2eb_rker_con_lat0_lon90.pdf} &
\hspace{\kernelfigspace}\includegraphics[width=\kernelfigwidth]{new_kernel/qu2eb_iker_con_lat0_lon90.pdf} \\
&
\centering $ \textrm{Re} \left(\mathcal{M}_{G} \right) $ &
\centering $\textrm{Im} \left(\mathcal{M}_{G} \right) $ &
\centering $\textrm{Re}  \left(\mathcal{M}_{B}^* \right) $ &
\centering $\textrm{Im} \left(\mathcal{M}_{B}^* \right) $
\end{tabular}
\end{center}
  \caption{Real and imaginary parts of real space kernels for $Q/U$ to $E/B$ translation (and vice versa).  The function $\mathcal{M}_{G}$ is the Green's function (radiation kernel) that gives $E+iB$ for a delta function Stokes input $_{+2}X = Q+iU$ at the center.  On the other hand, $\mathcal{M}^*_{B}$ is the Green's function (radiation kernel) that gives $_{+2}X$ for a delta function scalar input $E+iB$ at the center.  The black circles denotes the position of the center around which the kernels have been evaluated while the black star marks the location of the North Pole. The four rows depict the kernels at different latitudes on the sphere.   The kernels are unchanged for center points at constant latitude.  $\mathcal{M}_G$ is invariant over the sphere because $E/B$ fields are coordinate independent.  In the convolution kernels, $\mathcal{M}_G$ and $\mathcal{M}_B$ switch roles.  The kernels have been evaluated with the band limit $\ell \in [2,192]$ and sampled at the HEALPix resolution parameter NSIDE=2048. Each panel is approximately $16^{\circ} \times 16^{\circ}$ in size with grid lines every 2 degrees. } \label{fig:vis_kernel} 
\end{figure}

We compute the Euler angles $(\alpha, \beta, \gamma)$ given the angular coordinates of any two Healpix pixels and use these to evaluate the convolution and radiation kernels. To provide an intuition for how these kernels vary as a function of position of the central pixel we depict in \fig{fig:vis_kernel} the respective kernels at a few different locations on the sphere.
While the kernels are evaluated in the band limit $\ell \in [2,192]$, for illustration these functions are sampled at a very high Healpix resolution parameter of NSIDE=2048. All the plots have been rotated such that the central location marked by the black circle are in the centre of the figure. The horizontal and vertical lines that pass through the central black circle mark the local latitude and longitude respectively.
